---
title: 'Drama Series Analysis'
fontsize: 9pt
output:
  
  html_document: default
  pdf_document:
   latex_engine: xelatex
editor_options:
  
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(out.height = "\\textheight",  out.width = "\\textwidth",fig.width=9, fig.height=5)
```

```{r include = FALSE}
library(tidyverse)
library(lubridate)
library(broom)

tv_ratings_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-01-08/IMDb_Economist_tv_ratings.csv")
  
tv_ratings_processed <- tv_ratings_raw %>%
  filter(!seasonNumber > 20)%>%
    mutate(Drama = str_detect(genres,"Drama"),
         Mystery = str_detect(genres,"Mystery"),
         SciFi = str_detect(genres,"Sci-Fi"),
         Adventure = str_detect(genres,"Adventure"),
         Crime = str_detect(genres,"Crime"),
         Action =str_detect(genres,"Action"),
         Fantasy = str_detect(genres,"Fantasy"),
         Romance = str_detect(genres,"Romance"),
         Comedy = str_detect(genres,"Comedy"),
         Biography = str_detect(genres,"Biography"),
         Horror = str_detect(genres, "Horror"),
         Thriller = str_detect(genres,"Thriller"),
         Family = str_detect(genres,"Family"),
         Sport = str_detect(genres,"Sport"),
         History = str_detect(genres,"History"),
         War = str_detect(genres,"War"),
         Western = str_detect(genres,"Western"),
         Animation = str_detect(genres,"Animation"),
         Music = str_detect(genres,"Music"),
         decade = 10 * floor(year(date)/10))

tv_ratings_gathered <- tv_ratings_processed%>%
  gather(type,value,-(titleId:genres),-decade)%>%
  filter(value == 1, type != "Drama")

```
###Keep in mind that the analysis below is only about TV series that has a "Drama" genre attribute attached to it. With that said lets dive in!

How are TV dramas doing over the last 28 years?.

Could be due to an increase in interest of dramas, or due to an increase of improved quality of series in general.

&nbsp; 

&nbsp; 



```{r echo = FALSE}
  
  tv_ratings_gathered%>%
  group_by(year = year(date))%>%
  summarize(avg_rating = mean(av_rating),n())%>%
  ggplot(aes(year,avg_rating))+
  geom_line(col = "light blue", size = 1.5)+
  geom_point()+
  labs(x = "",y = "Mean rating",
       title = "Average rating for all TV dramas over the years")
```
&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

Histogram shows a normal distribution with a skewness to the left tail, mean and median are close,mean is used in the analysis.

```{r echo=FALSE, warning= FALSE , message= FALSE}
tv_ratings_gathered%>%
ggplot(aes(av_rating))+
  geom_histogram()+
  geom_vline(xintercept =  8.075958,color = "red",lty = "dashed")+
  geom_vline(xintercept = 8.112,color = "white",lty = "dashed")+
  labs(x= "IMDb rating", y= "# of occurrence", 
       title = "Histogram of ratings")
```
&nbsp; 

How did our taste change ovetime based on Average rating?. 

Most genres show no significant change, except for animation which has seen a rise in interest from the beginning of this decade. Another notable mentions is, War movies being assosciated with drama is a somewhat new combination of genres.

&nbsp; 

```{r echo = FALSE, message = FALSE}
tv_ratings_gathered%>%
ggplot(aes(date, av_rating))+
  geom_line()+
  geom_smooth()+
  facet_wrap(~type)+
  labs(x= "Date",y ="Average rating",
       title = "Trend of genres overtime")

```
&nbsp; 

&nbsp; 

&nbsp; 

How many series is the industry releasing per genre overtime?
```{r echo = FALSE}
tv_ratings_gathered%>%
  count(type,date = year(date))%>%
  ggplot(aes(x = date, y= n,col = type))+
  geom_line()+
  facet_wrap(~type)+
  labs(x = "Date", y="# of series",title = "Number of series per genre being made overtime")

```

Lets look closer to the most occuring genres, Action, Crime and Fantasy show an increase of average rating overtime, while comedy and mystery show no significant change, and a slight decrease of average rating for romance.
```{r echo = FALSE}


tv_ratings_gathered%>%
  mutate(type = fct_lump(type,6))%>%
  group_by(type,decade)%>%
  summarize(mean = mean(av_rating),number = n())%>%
  arrange(desc(mean))%>%
  ggplot(aes(x = decade,y=mean, col = type))+
  geom_line()+
  labs(x = "",y= "Average rating" , col = "Genre",
       title = "How has our intrest shifted through the decades?")
  

```
&nbsp; 

Looking for a long Drama series?.

These are the top 7 Drama series with more than 3 seasons, Game of Thrones takes the throne!
```{r echo = FALSE}
tv_ratings_processed%>%
  group_by(title)%>%
  summarize(meanrating= mean(av_rating),
            number= n())%>%
  filter(number >= 3)%>%
  arrange(desc(meanrating))%>%
  head(7)%>%
  ggplot(aes(reorder(title,meanrating),meanrating,fill = title))+
  geom_col(show.legend = FALSE)+
  coord_flip(ylim=c(9.0,9.3))+
  labs(x="",y="Average rating per series",
       title = "Best rated series with 3 or more seasons")
```

###Which had the best rating per season?
```{r echo = FALSE}
tv_ratings_gathered%>%
  group_by(title,seasonNumber)%>%
  summarize(MaxRating = max(av_rating))%>%
  arrange(desc(MaxRating))%>%
  head(6)%>%
  ungroup()%>%
  mutate(title = paste0(title," - Season ",seasonNumber))%>%
  ggplot(aes(reorder(title,MaxRating),MaxRating,fill = title))+
  geom_col(show.legend = FALSE)+
  coord_flip(ylim=c(9.4,9.7))+
  labs(x= "", y = "Rating",
       title = "Best season of all time?")
              



```

&nbsp; 

#Show survival
##Was the increase in ratings contributed to the interest of viewers overtime, or was it contributed to  of experience of the producers on knowing when to cancel a bad series overtime?

Before 2000, what this graph shows is that a rating of  "7-7.5" before 2000 had more than 60% chance of being renewed compared to a 33% chance after 2000, this could be contributed to the increase in competitiveness of series after 2000.

We can see that a rating of 8 and more decreased the chance of a second season before 2000, while it increased after 2000, this could be the increase in expertise of the industry in knowing viewer preferences.

```{r include= FALSE}
show_survival <- tv_ratings_processed%>%
  filter(seasonNumber <= 2)%>%
  group_by(title)%>%
  mutate(date_first_season = min(date))%>%
  ungroup()%>%
  transmute(titleId,
            title,
            date_first_season,
            seasonNumber = paste0("season",seasonNumber),
            av_rating)%>%
  distinct(title,seasonNumber, .keep_all = TRUE)%>%
  spread(seasonNumber,av_rating)%>%
  filter(!is.na(season1))%>%
  filter(date_first_season <= "2017-01-01")

 
```


```{r echo = FALSE}
show_survival %>%
  group_by(season1_bin = cut(season1,c(0,7,7.5,8,8.5,10)),
           time_bin = ifelse(date_first_season <"2000-01-01","Before 2000","After 2000"))%>%
  summarize(had_second_season = mean(!is.na(season2)),number = n())%>%
ggplot(aes(season1_bin,had_second_season,group = time_bin,col = time_bin))+
  geom_line()+
  geom_point()+
  scale_y_continuous(labels = scales:::percent_format())+
  labs(x="Season 1 rating ranges", y = "Percentage of second season", col = "",
       title = "Percentage of a series being continued to season 2 based on season 1")
```
&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 

&nbsp; 
Building a logistic regression model and testing it with outside data.

The model shows what we expected, a higher rated series in 1990 had a lower chance of being continued to season 2 than in 2018, this emphasizes the increase in expertise of producers overtime.

```{r echo = FALSE}
new_data <- crossing(year = 1990:2018,
                     season1 = seq(6,9))

show_survival%>%
  mutate(year = year(date_first_season),
         had_second_season = !is.na(season2))%>%
  glm(had_second_season ~ season1 * year ,data = ., family = "binomial")%>%
  augment(newdata = new_data , type.predict = "response")%>%
  ggplot(aes(year, .fitted, color = factor(season1)))+
  geom_line()+
  scale_y_continuous(labels = scales:::percent_format())+
  labs(x ="Date",y = "Percentage of second season", col = "Rating",
       title = "Percentage of season 2 based on rating")
  


  
```

